---
title: "Course Capstone Project"
author: "Lytemar"
date: "October 9, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache=TRUE)
set.seed(85)
```

## Introduction

The data set that we will analyze is the opening response time (ms) of an electromechanical valve.
The valve controls the flow of several fluids where the opening and closing is actuated by a solenoid.
The shutoff seal is an elastomer that is attached to the end of the solenoid plunger.
In this project, we will determine how the seal vintage, fluid and applied coil voltage affects the valve's opening response time.

## Exploratory Data Analysis

The summary of the data set is

```{r read in data and summary}
library(rjags)
library(ggplot2)
df = read.csv('capstone_data_set.csv')
df = df[, -c(1,2)]
df$Voltage = as.factor(df$Voltage)
prob = mean(df$Response.Time_ms[df$Voltage==30] <= df$Response.Time_ms[df$Voltage==24])
data_resp_time = df$Response.Time_ms
summary(df)
```

where *Valve.Rev* indicates the revision of the valve design, *Fluid* indicates the different fluids as masked factors, *Time.Between.Actuations_hr* indicates the time between valve actuations in hours, *\*.Age* indicates the age of the indicated component, *voltage* is the applied DC solenoid coil voltage, *Seal.Type* indicates the types of seal materials as masked factors, and *Response.Time_ms* is the time to open the valve in ms.

### Plots

In the interest of report length, we only show the following density plot that illustrates the distribution of response time.
Additionally, box plots of response time versus the other variables would show how response time varies due to the other variables in the data set.
Here, we see that the distribution of response time is right-skewed, therefore we will apply a log transformation before fitting a Bayesian regression model.


```{r pairs plot of data}
#library(GGally)
#ggpairs(df[,c(1,2,6,9)], aes(color=Voltage, alpha=0.4))

p <- ggplot(data=df, aes(x=Response.Time_ms))
p <- p + geom_density()
print(p)

#p <- ggplot(data=df, aes(x=Response.Time_ms, color=Voltage, group=Voltage))
#p <- p + geom_density()
#print(p)

#p <- ggplot(df, aes(Voltage, Response.Time_ms))
#p <- p + geom_boxplot(outlier.colour = "red", outlier.shape = 1, aes(color=Fluid))
#p <- p + + geom_jitter(width=0.2)
#print(p)
```

## Modeling

Here we will fit a Bayesian linear model from which we will see which predictors affect response time.
The coefficients of the predictors are

- b[1]: intercept,
- b[2]: is.new.vlv,
- b[3]: Fluid,
- b[4]: Time.Between.Actuations_ms,
- b[5]: is.new.plunger,
- b[6]: is.new.coil,
- b[7]: is.24v,
- b[8]: is.seal_A,
- b[9]: is.new.seal,
- b[10]: Fluid*is.24v,
- b[11]: is.new.seal*is.24v, and
- b[12]: is.new.vlv*is.24v.

Finally, we will compute the probability that the response time is less than 100 ms given Voltage from our MCMC results.

```{r jags model}

df$logresponse = log(df$Response.Time_ms)
df$Response.Time_ms <- NULL

# Convert all factors to numeric
#indx <-sapply(df, is.factor)
#df[indx] <- lapply(df[indx], function(x) as.numeric(x))
df$is.new.vlv = as.numeric(df$Valve.Rev=='New')
df$is.new.plunger = as.numeric(df$Plunger.Age=='New')
df$is.new.coil = as.numeric(df$Coil.Age=='New')
df$is.24v = as.numeric(df$Voltage=='24')
df$is.seal_A = as.numeric(df$Seal.Type=='A')
df$is.new.seal = as.numeric(df$Seal.Age=='New')
df$Fluid = as.numeric(df$Fluid)

n.interaction.terms = 3
n.chains = 3
n.burn.iter = 5000
n.iter = 100000

# Make model matrix for interactions
#M = model.matrix(logresponse ~ .^2, df)

mod_string = " model {
  for (i in 1:n) {
    y[i] ~ dnorm(mu[i], prec)
    mu[i] = b[1]+b[2]*is.new.vlv[i]+b[3]*fluid[i]+b[4]*time[i]+b[5]*is.new.plunger[i]+b[6]*is.new.coil[i]+b[7]*is.24v[i]+b[8]*is.seal_A[i]+b[9]*is.new.seal[i]+b[10]*is.24v[i]*fluid[i]+b[11]*is.24v[i]*is.new.seal[i]+b[12]*is.24v[i]*is.new.vlv[i]
  }
  
  for (j in 1:m) {
    b[j] ~ dnorm(0.0, 1.0/1.0e2)
  }
  
  prec ~ dgamma(5/2.0, 2*10.0/2.0)
  sig2 = 1.0/prec
  sig = sqrt(sig2)
} "

data_jags = list(n=nrow(df), 
                 m=9+n.interaction.terms, 
                 y=df$logresponse,
                 is.new.vlv=df$is.new.vlv,
                 fluid = df$Fluid,
                 time = df$Time.Between.Actuations_hr,
                 is.new.plunger = df$is.new.plunger,
                 is.new.coil = df$is.new.coil,
                 is.24v = df$is.24v,
                 is.seal_A = df$is.seal_A,
                 is.new.seal = df$is.new.seal
                 )
params = c("b", "sig2")
mod = jags.model(textConnection(mod_string), data=data_jags, n.chains=n.chains)
update(mod, n.burn.iter) # burn-in
mod_sim = coda.samples(model=mod,
                       variable.names = params,
                       n.iter = n.iter)
mod_csim = do.call(rbind, mod_sim) # combine multiple chains
#gelman.diag(mod_sim)
#autocorr.diag(mod_sim)
#autocorr.plot(mod_sim)
#effectiveSize(mod_sim)
#summary(mod_sim)

```

Using normal likelihood, with normal priors on the coefficients and inverse gamma prior for the variance, with `r as.integer(n.burn.iter)` burn-in iterations, and `r as.integer(n.iter)` total iterations for the `r as.integer(n.chains)` chains, we get the following autocorrelation diagnostics

```{r}
autocorr.diag(mod_sim)
```

Here, we see that not all of the parameters have converged, giving the following effective sample sizes
```{r}
effectiveSize(mod_sim)
```

However, we are interested in the means of the coefficients which are
```{r}
pm_params = colMeans(mod_csim)
pm_params
```

The plot of the residuals shows that we have some dependency on the variance with respect to the data point.
This would need to be remedied by a different transformation on the data.

```{r plot of residuals, fig.height=4, fig.width=6}
X = as.matrix(cbind(1, df[,c(10,2,3,11,12,13,14,15)],
          df$Fluid*df$is.24v,
          df$is.new.seal*df$is.24v,
          df$is.24v*df$is.new.vlv))
last.pm.params.col=9+n.interaction.terms+1
log_yhat = X %*% pm_params[seq(1,9+n.interaction.terms)]
resid = data_jags$y - log_yhat
plot(resid)
```

## Results
The following plot superinposes the log response time distributions of the data and resulting model.

```{r density plots, fig.height=3, fig.width=6}
modeled_data = data.frame(Response.Time_ms=exp(log_yhat), vec='model')
observed_data = data.frame(Response.Time_ms=data_resp_time, vec='observed')
data = rbind(modeled_data, observed_data)
p <- ggplot(data, aes(Response.Time_ms, group=vec, col=vec)) + geom_density()
print(p)
```

Finally, using the modeled posterior distribution, the probability that the response time given 30 volts is less than given 24 volts.
```{r}
n_sum = nrow(mod_csim)
X.24 = X[X[,6]==1,]
X.30 = X[X[,6]==0,]
mod_csim = mod_csim[, seq(1,9+n.interaction.terms)]
prob
```


